---
title: "An Introduction to ADMerge"
author: "Shijia Zhang, Xinyu Guo, Zheyu Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Introduction to MyPackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Introduction
============

ADMerge is tailored for aggregating Alzheimer’s Disease (AD) biomarker data. It equips users with functionality to integrate data from diverse AD-related sources into a cohesive analytical dataset. Additionally, the package includes built-in functions that assist in summarizing the input and merged data, as well as in visualizing the analytical dataset for enhanced insights. 



```{r setup, include = FALSE}
library(devtools)
install_github("Shijia1997/ADMerge")
library(ADMerge)

```

# Generating dictionary for data used to be merged example Biocard data

```{r}
Biocard_path = system.file("data/BIOCARD August 2021 internal", package = "ADMerge")



src_table = get_src_table(path =sample_path,ID_pattern = "ID",DATE_pattern = "VIS")


```

