---
title: "test_plot_github"
author: "Shijia Zhang"
date: "2024-04-09"
output: html_document
---

```{r}
library(devtools)
install_github("Shijia1997/ADMerge",upgrade = "never")
library(ADMerge)
library(tidyr)
```

```{r}
data_path = "ADNI data"

scr_table = get_src_table(path = data_path, ID_pattern = "ID", DATE_pattern = "VIS|DATE",non_longitudinal_list = c("DESIKANLAB_14Apr2024"), ID_usr_list = c("RID"), DATE_usr_list = c("VISCODE"))

print(scr_table)



```

Each file identified as longitudinal will be join with the their "ID_for_merge" as y-axis and "DATE_for_merge" as x-axis for visualization. The DATE_for_merge from different files are integrated to the same timeline

By pass the source table into the "plot.files" function, we are able to visualize the distribution.

The user are able to hang over their mouses to points in the plot to highlight the all the scatter belongs to that group (file) while the users can also use the dropdown bar to visualize each file indivudally

```{r}
plot.files(path = data_path, dict_src = scr_table)
```


```{r}
library(data.table)

item = fread("/Users/shijia.zhang/Desktop/ADNI/ADMerge/data/ADNI data/ITEM.csv")

item %>% select(RID,VISCODE)

item

ADS <- fread("/Users/shijia.zhang/Downloads/ADSXLIST_13Apr2024.csv")

ADS %>% select(RID,VISCODE,Phase) %>% filter(RID == "667")
```

```{r}
upen_plasma <- fread("/Users/shijia.zhang/Desktop/ADNI/ADMerge/data/ADNI data/UPENNPLASMA_02Feb2024.csv")

upen_plasma %>% distinct(RID,VISCODE)

desk <- fread("/Users/shijia.zhang/Downloads/DESIKANLAB_14Apr2024.csv")

desk %>% distinct(RID)
```

