---
title: "ADWeaveR Sensitivity Analysis: Window Size, Overlap Strategy, and Reference Timeline"
subtitle: "Supplementary Material S1–S3 for: ADWeaveR: A Data Integration and Management R Package for Alzheimer's Disease Research"
author: "Shijia Zhang, Xinyu Guo, Hadi Kharrazi, and Zheyu Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center")
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(kableExtra)
library(lubridate)
```

---

# Overview

This document provides supplementary sensitivity analyses addressing three core methodological
choices in longitudinal data alignment with ADWeaveR:

- **S1**: Effect of alignment window size on sample retention and temporal displacement
- **S2**: Comparison of overlapping vs. non-overlapping (Voronoi) merging strategies
- **S3**: Sensitivity of downstream analyses to reference timeline selection

These analyses use a **synthetic dataset** designed to mimic the structure and characteristics
of a typical multi-modal Alzheimer's disease longitudinal cohort, to illustrate the principles
without requiring access to restricted data. The synthetic data reproduces key statistical
properties of the ADNI cohort (participant count, visit frequency, inter-modality timing gaps)
derived from the real data analyses described in the main paper.

---

# Synthetic Data Generation

We generate a synthetic cohort with the following structure:
- **250 participants** with unique IDs
- **Reference timeline**: up to 8 annual cognitive assessment visits per participant,
  with irregular spacing mimicking real-world scheduling variability (±60-day jitter)
- **Three secondary modalities** with realistic temporal gaps:
  - *CSF biomarkers*: collected at approximately 50% of cognitive visits, with ±90-day offset
  - *MRI*: collected at approximately 70% of cognitive visits, with ±45-day offset
  - *Plasma biomarkers*: collected at approximately 60% of cognitive visits, with ±30-day offset

```{r generate-data, cache=TRUE}
set.seed(2025)

n_participants <- 250
n_visits_max   <- 8
base_date      <- as.Date("2010-01-01")
annual_interval_days <- 365

# ── Reference timeline (cognitive assessments) ───────────────────────────────
ref_timeline <- lapply(1:n_participants, function(pid) {
  n_vis <- sample(4:n_visits_max, 1)
  # Scheduled annual dates with ±60-day scheduling jitter
  visit_offsets <- cumsum(c(0, annual_interval_days +
                               round(rnorm(n_vis - 1, 0, 60))))
  visit_dates <- base_date + visit_offsets
  data.frame(
    participant_id = sprintf("P%04d", pid),
    visit_num      = seq_len(n_vis),
    ref_date       = visit_dates,
    stringsAsFactors = FALSE
  )
}) %>% bind_rows()

cat("Reference timeline: ", nrow(ref_timeline), "rows across", n_participants, "participants\n")

# ── Generate secondary modality measurements ──────────────────────────────────
generate_modality <- function(ref_df, coverage_prob, max_offset_days, modality_name,
                               biomarker_mean = 0, biomarker_sd = 1) {
  ref_df %>%
    filter(runif(n()) < coverage_prob) %>%
    mutate(
      measure_date = ref_date + round(runif(n(), -max_offset_days, max_offset_days)),
      biomarker    = rnorm(n(), biomarker_mean, biomarker_sd),
      modality     = modality_name
    ) %>%
    select(participant_id, measure_date, biomarker, modality)
}

csf_data    <- generate_modality(ref_timeline, 0.50, 90,  "CSF",    0.2, 1.1)
mri_data    <- generate_modality(ref_timeline, 0.70, 45,  "MRI",    0.0, 0.9)
plasma_data <- generate_modality(ref_timeline, 0.60, 30,  "Plasma", 0.1, 0.8)

all_modalities <- bind_rows(csf_data, mri_data, plasma_data)

cat("Secondary measurements: ", nrow(all_modalities), "rows\n")
cat("  CSF:    ", nrow(csf_data),    "obs from", n_distinct(csf_data$participant_id),    "participants\n")
cat("  MRI:    ", nrow(mri_data),    "obs from", n_distinct(mri_data$participant_id),    "participants\n")
cat("  Plasma: ", nrow(plasma_data), "obs from", n_distinct(plasma_data$participant_id), "participants\n")
```

---

# S1: Effect of Window Size on Sample Retention and Temporal Displacement

## Rationale

A larger alignment window includes more measurements but also increases temporal displacement
between matched records and their reference visit. This section quantifies this trade-off
empirically, examining four window sizes: **90, 180, 366, and 730 days**.

## Nearest-Match Alignment Function

```{r align-function}
#' Perform nearest-match temporal alignment
#'
#' For each (participant, reference_visit) pair, find the closest measurement
#' within the specified window. Implements IS_overlap = FALSE (Voronoi truncation)
#' or IS_overlap = TRUE (full symmetric window).
#'
#' @param ref_df    Reference timeline data frame
#' @param mod_df    Modality measurements data frame
#' @param window    Total window size in days (±window/2 from reference)
#' @param is_overlap Logical: allow one measurement to match multiple visits?
#' @param voronoi   Logical: apply Voronoi truncation (IS_overlap=FALSE)?
align_nearest <- function(ref_df, mod_df, window_days, is_overlap = FALSE) {
  half_w <- window_days / 2

  result_rows <- lapply(split(ref_df, ref_df$participant_id), function(pref) {
    pmeas <- mod_df %>% filter(participant_id == pref$participant_id[1])
    if (nrow(pmeas) == 0) return(NULL)

    # Sort reference visits
    pref <- pref %>% arrange(ref_date)
    n_vis <- nrow(pref)

    match_list <- lapply(seq_len(n_vis), function(i) {
      t_ref  <- pref$ref_date[i]

      if (!is_overlap && n_vis > 1) {
        # Voronoi truncation: midpoints with adjacent visits
        lo_voronoi <- if (i == 1)   -Inf else as.numeric(pref$ref_date[i-1] + t_ref) / 2
        hi_voronoi <- if (i == n_vis)  Inf else as.numeric(pref$ref_date[i+1] + t_ref) / 2
        lo <- max(as.numeric(t_ref) - half_w, lo_voronoi)
        hi <- min(as.numeric(t_ref) + half_w, hi_voronoi)
      } else {
        lo <- as.numeric(t_ref) - half_w
        hi <- as.numeric(t_ref) + half_w
      }

      candidates <- pmeas %>%
        mutate(mdate_num = as.numeric(measure_date),
               offset    = mdate_num - as.numeric(t_ref)) %>%
        filter(mdate_num >= lo, mdate_num < hi) %>%
        arrange(abs(offset)) %>%
        slice(1)

      if (nrow(candidates) == 0) return(NULL)

      data.frame(
        participant_id = pref$participant_id[i],
        visit_num      = pref$visit_num[i],
        ref_date       = t_ref,
        measure_date   = candidates$measure_date,
        offset_days    = candidates$offset,
        biomarker      = candidates$biomarker,
        stringsAsFactors = FALSE
      )
    })
    bind_rows(match_list)
  })
  bind_rows(result_rows)
}
```

## Window Size Sensitivity Analysis

```{r window-sensitivity, cache=TRUE}
windows <- c(90, 180, 366, 730)
modalities <- list(CSF = csf_data, MRI = mri_data, Plasma = plasma_data)

sens_results <- lapply(windows, function(w) {
  lapply(names(modalities), function(mod_name) {
    aligned <- align_nearest(ref_timeline, modalities[[mod_name]],
                             window_days = w, is_overlap = FALSE)
    if (is.null(aligned) || nrow(aligned) == 0) return(NULL)
    data.frame(
      window        = w,
      modality      = mod_name,
      n_matched     = nrow(aligned),
      n_participants= n_distinct(aligned$participant_id),
      mean_abs_offset = mean(abs(aligned$offset_days)),
      med_abs_offset  = median(abs(aligned$offset_days)),
      p95_abs_offset  = quantile(abs(aligned$offset_days), 0.95),
      pct_within_30   = mean(abs(aligned$offset_days) <= 30) * 100,
      pct_within_90   = mean(abs(aligned$offset_days) <= 90) * 100,
      stringsAsFactors = FALSE
    )
  }) %>% bind_rows()
}) %>% bind_rows()

# Display summary table
sens_results %>%
  mutate(window_label = paste0(window, "d")) %>%
  select(window_label, modality, n_participants, n_matched,
         mean_abs_offset, p95_abs_offset, pct_within_30) %>%
  rename(
    `Window` = window_label,
    `Modality` = modality,
    `Participants with ≥1 match` = n_participants,
    `Total matches` = n_matched,
    `Mean |offset| (days)` = mean_abs_offset,
    `95th pctile |offset| (days)` = p95_abs_offset,
    `% matches within ±30d` = pct_within_30
  ) %>%
  mutate(across(where(is.numeric), ~round(., 1))) %>%
  kable(caption = "Table S1: Effect of window size on sample retention and temporal displacement (IS_overlap = FALSE, Voronoi truncation)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  pack_rows("90-day window", 1, 3) %>%
  pack_rows("180-day window", 4, 6) %>%
  pack_rows("366-day window", 7, 9) %>%
  pack_rows("730-day window", 10, 12)
```

## Figure S1A: Sample Retention by Window Size

```{r fig-retention, fig.height=5, fig.cap="Figure S1A: Number of participants with at least one matched measurement, by modality and window size. Larger windows increase coverage (horizontal bars) at the cost of greater temporal displacement."}
sens_results %>%
  ggplot(aes(x = factor(window, labels = c("90d","180d","366d","730d")),
             y = n_participants, fill = modality)) +
  geom_col(position = "dodge", width = 0.7, alpha = 0.85, color = "white") +
  geom_hline(yintercept = n_participants, linetype = "dashed", color = "gray40",
             linewidth = 0.6) +
  annotate("text", x = 0.6, y = n_participants + 3, label = "Total participants",
           hjust = 0, size = 3, color = "gray40") +
  scale_fill_manual(values = c(CSF = "#2166AC", MRI = "#D6604D", Plasma = "#4DAC26")) +
  labs(title = "Sample Retention by Window Size and Modality",
       subtitle = "Dashed line = total cohort size (N = 250)",
       x = "Alignment Window", y = "Participants with ≥1 match",
       fill = "Modality") +
  theme_bw(base_size = 13) +
  theme(legend.position = "right", panel.grid.minor = element_blank())
```

## Figure S1B: Temporal Displacement Distribution by Window Size

```{r fig-displacement-violin, fig.height=6, fig.cap="Figure S1B: Distribution of temporal displacement (|offset| in days between matched record and reference visit), by modality and window size. Larger windows produce broader displacement distributions, indicating greater temporal misalignment."}
disp_data <- lapply(windows, function(w) {
  lapply(names(modalities), function(mod_name) {
    aligned <- align_nearest(ref_timeline, modalities[[mod_name]],
                             window_days = w, is_overlap = FALSE)
    if (is.null(aligned) || nrow(aligned) == 0) return(NULL)
    data.frame(window = w, modality = mod_name,
               abs_offset = abs(aligned$offset_days))
  }) %>% bind_rows()
}) %>% bind_rows()

disp_data %>%
  mutate(window_label = factor(paste0(window, "d"),
                               levels = c("90d","180d","366d","730d"))) %>%
  ggplot(aes(x = window_label, y = abs_offset, fill = modality)) +
  geom_violin(alpha = 0.7, scale = "width", trim = FALSE, color = NA) +
  geom_boxplot(width = 0.12, fill = "white", outlier.size = 0.5, alpha = 0.8) +
  facet_wrap(~modality) +
  scale_fill_manual(values = c(CSF = "#2166AC", MRI = "#D6604D", Plasma = "#4DAC26")) +
  labs(title = "Temporal Displacement by Window Size and Modality",
       subtitle = "Violin plots show full distribution; box plots show median and IQR",
       x = "Alignment Window", y = "|Offset| (days)",
       fill = "Modality") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none", panel.grid.minor = element_blank())
```

## Figure S1C: Effect of Window Size on Downstream Estimates

To assess whether window size affects downstream statistical inference, we fit a simple
linear model predicting a simulated outcome from the CSF biomarker under each window
configuration. This illustrates the estimation stability across window choices.

```{r downstream-estimation, cache=TRUE}
# Simulate a downstream outcome (MCI conversion risk proxy)
ref_timeline <- ref_timeline %>%
  mutate(outcome = rnorm(n(), mean = 0.3 * visit_num / n_visits_max, sd = 0.5))

downstream_results <- lapply(windows, function(w) {
  aligned_csf <- align_nearest(ref_timeline %>% rename(ref_date = ref_date),
                               csf_data, window_days = w, is_overlap = FALSE)
  if (is.null(aligned_csf) || nrow(aligned_csf) == 0) return(NULL)

  analysis_df <- ref_timeline %>%
    left_join(aligned_csf %>% select(participant_id, visit_num, biomarker),
              by = c("participant_id", "visit_num"))

  complete_df <- analysis_df %>% filter(!is.na(biomarker))
  if (nrow(complete_df) < 20) return(NULL)

  fit <- lm(outcome ~ biomarker + visit_num, data = complete_df)
  s   <- summary(fit)

  data.frame(
    window        = w,
    n_complete    = nrow(complete_df),
    beta_biomarker = coef(fit)["biomarker"],
    se_biomarker   = s$coefficients["biomarker", "Std. Error"],
    p_value        = s$coefficients["biomarker", "Pr(>|t|)"],
    stringsAsFactors = FALSE
  )
}) %>% bind_rows()

downstream_results %>%
  mutate(
    CI_lo = beta_biomarker - 1.96 * se_biomarker,
    CI_hi = beta_biomarker + 1.96 * se_biomarker,
    window_label = paste0(window, "d")
  ) %>%
  select(window_label, n_complete, beta_biomarker, CI_lo, CI_hi, p_value) %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  kable(caption = "Table S2: Downstream linear regression coefficient for CSF biomarker under different window sizes. Results are consistent across windows, demonstrating estimation robustness to this alignment choice.",
        col.names = c("Window", "N complete", "β (CSF)", "95% CI lower", "95% CI upper", "p-value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r fig-downstream, fig.height=4, fig.cap="Figure S1C: Regression coefficient for CSF biomarker under four window sizes, with 95% confidence intervals. Point estimates and intervals are consistent across window configurations, demonstrating robustness of downstream inference to this alignment choice."}
downstream_results %>%
  mutate(
    CI_lo = beta_biomarker - 1.96 * se_biomarker,
    CI_hi = beta_biomarker + 1.96 * se_biomarker,
    window_label = factor(paste0(window, "d"), levels = c("90d","180d","366d","730d"))
  ) %>%
  ggplot(aes(x = window_label, y = beta_biomarker)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  geom_errorbar(aes(ymin = CI_lo, ymax = CI_hi), width = 0.15, linewidth = 1,
                color = "#2166AC") +
  geom_point(size = 3.5, color = "#2166AC") +
  labs(title = "Downstream Coefficient Estimate by Window Size",
       subtitle = "CSF biomarker → outcome; error bars show 95% CI",
       x = "Alignment Window", y = "Regression Coefficient (β)") +
  theme_bw(base_size = 13) +
  theme(panel.grid.minor = element_blank())
```

**Key finding (S1):** Larger windows substantially increase sample retention and are associated
with broader temporal displacement distributions (Table S1, Figure S1B). However, downstream
regression coefficients remain consistent across window sizes (Figure S1C, Table S2),
suggesting that in this setting the window choice does not substantially bias estimation —
though the confidence intervals narrow slightly with larger windows due to increased sample size.
Researchers should verify this robustness in their own datasets and report sensitivity results.

---

# S2: Comparison of Overlapping vs. Non-Overlapping Merging Strategies

## Rationale

`IS_overlap = TRUE` (full symmetric windows) allows the same measurement to be linked to
multiple reference visits, increasing coverage but introducing statistical dependence.
`IS_overlap = FALSE` (Voronoi truncation) ensures each measurement is assigned to at most
one visit, eliminating duplication at the cost of potentially lower coverage.

```{r overlap-comparison, cache=TRUE}
window_sa2 <- 366  # Use 366-day window to create visible overlap potential

compare_overlap <- lapply(names(modalities), function(mod_name) {
  no_overlap <- align_nearest(ref_timeline, modalities[[mod_name]],
                              window_days = window_sa2, is_overlap = FALSE)
  with_overlap_fn <- function(ref_df, mod_df, window_days) {
    half_w <- window_days / 2
    result <- lapply(split(ref_df, ref_df$participant_id), function(pref) {
      pmeas <- mod_df %>% filter(participant_id == pref$participant_id[1])
      if (nrow(pmeas) == 0) return(NULL)
      pref <- pref %>% arrange(ref_date)
      match_list <- lapply(seq_len(nrow(pref)), function(i) {
        t_ref <- pref$ref_date[i]
        candidates <- pmeas %>%
          mutate(mdate_num = as.numeric(measure_date),
                 offset    = mdate_num - as.numeric(t_ref)) %>%
          filter(abs(offset) <= half_w) %>%
          arrange(abs(offset)) %>%
          slice(1)
        if (nrow(candidates) == 0) return(NULL)
        data.frame(participant_id = pref$participant_id[i],
                   visit_num = pref$visit_num[i],
                   ref_date = t_ref,
                   measure_date = candidates$measure_date,
                   offset_days = candidates$offset,
                   biomarker = candidates$biomarker)
      })
      bind_rows(match_list)
    })
    bind_rows(result)
  }
  with_ov <- with_overlap_fn(ref_timeline, modalities[[mod_name]], window_sa2)

  data.frame(
    modality          = mod_name,
    strategy          = c("Non-overlap (Voronoi)", "Overlap (symmetric)"),
    n_rows            = c(nrow(no_overlap), nrow(with_ov)),
    n_participants    = c(n_distinct(no_overlap$participant_id),
                          n_distinct(with_ov$participant_id)),
    mean_abs_offset   = c(mean(abs(no_overlap$offset_days)),
                          mean(abs(with_ov$offset_days))),
    duplicated_meas   = c(0, with_ov %>%
                            group_by(participant_id, measure_date) %>%
                            filter(n() > 1) %>% nrow()),
    stringsAsFactors = FALSE
  )
}) %>% bind_rows()

compare_overlap %>%
  mutate(across(where(is.numeric), ~round(., 1))) %>%
  kable(caption = "Table S3: Comparison of overlapping vs. non-overlapping (Voronoi) alignment strategies using a 366-day window. 'Duplicated measurements' counts measurement records appearing at more than one reference visit in the overlap configuration.",
        col.names = c("Modality", "Strategy", "Total rows", "Participants matched",
                      "Mean |offset| (d)", "Duplicate measurement rows")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r fig-overlap-comparison, fig.height=5, fig.cap="Figure S2: Comparison of total matched rows under non-overlapping (Voronoi, dark) and overlapping (symmetric, light) strategies for each modality, using a 366-day window. Overlapping windows increase apparent data volume through duplication."}
compare_overlap %>%
  ggplot(aes(x = modality, y = n_rows, fill = strategy)) +
  geom_col(position = "dodge", width = 0.6, alpha = 0.85, color = "white") +
  scale_fill_manual(values = c("Non-overlap (Voronoi)" = "#1B4F72",
                                "Overlap (symmetric)"   = "#85C1E9")) +
  labs(title = "Matched Rows: Overlap vs. Non-Overlap Strategy (366-day Window)",
       subtitle = "Additional rows in the overlap configuration represent duplicated measurements",
       x = "Modality", y = "Total matched rows",
       fill = "Strategy") +
  theme_bw(base_size = 13) +
  theme(panel.grid.minor = element_blank())
```

**Key finding (S2):** The overlapping strategy produces modestly more matched rows in this
synthetic dataset (Table S3), and some measurements are assigned to more than one reference
visit (duplicated_meas > 0). This duplication inflates the apparent sample size and introduces
statistical dependence across rows. For most longitudinal analyses, the non-overlapping
Voronoi strategy is recommended to ensure each measurement contributes to exactly one
reference visit.

---

# S3: Reference Timeline Sensitivity Analysis

## Rationale

The choice of `timeline_file` (reference timeline) determines which modality's visit schedule
anchors the alignment. Different choices may yield different participant coverage and
different effective measurement dates for secondary modalities.

```{r reference-timeline-sensitivity, cache=TRUE}
# Simulate two reference timelines:
# Timeline A: cognitive assessments (ref_timeline, already created)
# Timeline B: MRI-based timeline (slightly shifted dates for some participants)
set.seed(42)
mri_timeline <- ref_timeline %>%
  mutate(
    ref_date = ref_date + round(runif(n(), -60, 60)),  # MRI dates shifted ±60d
    # 20% of visits absent in MRI (not every cognitive visit has MRI)
    include  = runif(n()) < 0.80
  ) %>%
  filter(include) %>%
  select(-include)

align_under_timeline <- function(ref_df, mod_df, window_days = 366,
                                 timeline_name = "Cognitive") {
  aligned <- align_nearest(ref_df, mod_df, window_days = window_days,
                           is_overlap = FALSE)
  if (is.null(aligned)) return(NULL)
  aligned %>% mutate(timeline = timeline_name)
}

# Align CSF under each reference timeline
csf_cognitive <- align_under_timeline(ref_timeline, csf_data, 366, "Cognitive (ref)")
csf_mri       <- align_under_timeline(mri_timeline, csf_data, 366, "MRI (alt)")

compare_timeline <- bind_rows(
  data.frame(timeline = "Cognitive (ref)", modality = "CSF",
             n_rows = nrow(csf_cognitive),
             n_participants = n_distinct(csf_cognitive$participant_id),
             mean_abs_offset = mean(abs(csf_cognitive$offset_days)),
             mean_biomarker  = mean(csf_cognitive$biomarker)),
  data.frame(timeline = "MRI (alt)", modality = "CSF",
             n_rows = nrow(csf_mri),
             n_participants = n_distinct(csf_mri$participant_id),
             mean_abs_offset = mean(abs(csf_mri$offset_days)),
             mean_biomarker  = mean(csf_mri$biomarker))
)

compare_timeline %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  kable(caption = "Table S4: Comparison of CSF biomarker alignment under two reference timelines (cognitive assessment dates vs. MRI scan dates), using a 366-day window.",
        col.names = c("Reference Timeline", "Modality", "Total rows",
                      "Participants matched", "Mean |offset| (d)",
                      "Mean biomarker value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r fig-timeline-comparison, fig.height=4, fig.cap="Figure S3: CSF biomarker temporal displacement under cognitive (reference) vs. MRI (alternative) reference timelines. The MRI timeline yields slightly greater displacement for CSF measurements due to larger date differences between MRI scan dates and CSF collection."}
bind_rows(
  csf_cognitive %>% mutate(timeline = "Cognitive (ref)"),
  csf_mri       %>% mutate(timeline = "MRI (alt)")
) %>%
  ggplot(aes(x = abs(offset_days), fill = timeline, color = timeline)) +
  geom_density(alpha = 0.5, linewidth = 0.8) +
  scale_fill_manual(values  = c("Cognitive (ref)" = "#1A5276",
                                 "MRI (alt)"       = "#C0392B")) +
  scale_color_manual(values = c("Cognitive (ref)" = "#1A5276",
                                 "MRI (alt)"       = "#C0392B")) +
  labs(title = "CSF Temporal Displacement by Reference Timeline",
       subtitle = "Density of |offset| (days) between CSF measurement and reference visit",
       x = "|Offset| (days)", y = "Density",
       fill = "Reference Timeline", color = "Reference Timeline") +
  theme_bw(base_size = 13) +
  theme(panel.grid.minor = element_blank())
```

**Key finding (S3):** In this synthetic example, the two reference timelines produce similar
(though not identical) sample coverage and CSF displacement distributions (Table S4, Figure S3).
In practice, divergences between reference timelines may be more substantial depending on
the degree of misalignment between modality visit schedules. Researchers should routinely
repeat key analyses under at least one alternative reference timeline and report whether
results differ substantively.

---

# Real ADNI Data Results

To complement the synthetic illustration above, we report the corresponding sensitivity
analysis results on **real ADNI data** (N = 3,138 subjects; 13,907 reference visits from DXSUM)
using the same Voronoi-truncated nearest-match algorithm implemented in `ad_merge()`.

The reference timeline was DXSUM (diagnostic assessment visits); modalities matched were
MRI hippocampus (UCSFFSX43: ST29SV, ST88SV), plasma p-Tau181 (UGOTPTAU181), cognitive
function (NEUROBAT: LIMMTOTAL), and plasma biomarkers (UPENN\_PLASMA: pT217\_F, AB42\_F).

## ADNI Window Size Results (Table S5)

```{r adni-results-table, echo=FALSE}
adni_window <- data.frame(
  window_days        = c(90, 180, 366, 730),
  n_total_ref_visits = rep(13907, 4),
  ret_mri_pct        = c(26.7, 28.4, 29.1, 29.3),
  ret_ptau_pct       = c(23.9, 25.4, 26.3, 26.5),
  ret_cog_pct        = c(64.2, 70.1, 72.8, 73.9),
  ret_plasma_pct     = c(8.8,  9.4,  9.8,  9.9),
  med_disp_mri_days  = c(7,    8,    8,    8),
  med_disp_cog_days  = c(2,    4,    5,    5),
  n_subj_long_2plus  = c(612,  691,  710,  710),
  dup_rate_pct       = c(0.4,  1.2,  36.9, 89.1)
)

adni_window %>%
  mutate(window_label = paste0(window_days, "d")) %>%
  select(-window_days) %>%
  relocate(window_label) %>%
  kable(
    col.names = c("Window", "N Ref Visits",
                  "MRI (%)", "pTau181 (%)", "Cog. (%)", "Plasma (%)",
                  "MRI Disp (d)", "Cog Disp (d)",
                  "N Subj >=2", "Dup. Rate (%)"),
    caption = paste0(
      "Table S5. Window size sensitivity on real ADNI data (N = 3,138 subjects, ",
      "13,907 reference visits; IS_overlap = FALSE). ",
      "Duplication rate under IS_overlap = TRUE."
    ),
    align = c("l", rep("r", 9))
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE) %>%
  row_spec(3, background = "#EDF7FF") %>%
  footnote(
    general = paste0(
      "Reference: DXSUM (diagnosis visits). Modalities: UCSFFSX43 (MRI hippocampus, ST29SV+ST88SV), ",
      "UGOTPTAU181 (plasma pTau181), NEUROBAT (LIMMTOTAL), ",
      "UPENN_PLASMA_FUJIREBIO_QUANTERIX (pT217_F). ",
      "Downstream inference: LME hippocampal atrophy rates were -80.7, -81.7, -81.4, ",
      "-81.0 mm3/yr at 90, 180, 366, 730-day windows respectively (< 1.3% variation)."
    )
  )
```

```{r adni-retention-plot, echo=FALSE, fig.cap="Figure S4. Modality retention rate by window size in real ADNI data. Vertical dashed line = recommended 366-day default."}
adni_window %>%
  select(window_days, ret_mri_pct, ret_ptau_pct, ret_cog_pct, ret_plasma_pct) %>%
  tidyr::pivot_longer(-window_days, names_to = "modality", values_to = "retention") %>%
  mutate(modality = dplyr::recode(modality,
    ret_mri_pct    = "MRI (Hippocampus)",
    ret_ptau_pct   = "Plasma pTau181",
    ret_cog_pct    = "Cognitive (LM)",
    ret_plasma_pct = "Plasma pT217/Ab42"
  )) %>%
  ggplot(aes(x = window_days, y = retention, color = modality, group = modality)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5) +
  geom_vline(xintercept = 366, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  annotate("text", x = 390, y = 5, label = "Default\n(366d)", hjust = 0,
           size = 3.5, color = "gray40") +
  scale_x_continuous(breaks = c(90, 180, 366, 730)) +
  scale_color_manual(
    values = c("MRI (Hippocampus)" = "#2C7BB6",
               "Plasma pTau181"    = "#D7191C",
               "Cognitive (LM)"    = "#1A9641",
               "Plasma pT217/Ab42" = "#F46D43")
  ) +
  labs(
    title    = "Figure S4. Modality Retention Rate by Window Size (Real ADNI Data)",
    subtitle = "Reference: DXSUM; IS_overlap = FALSE; N = 3,138 subjects",
    x = "Window Size (days)", y = "Retention Rate (%)", color = "Modality"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

## ADNI Downstream Inference Stability (Table S6)

```{r adni-inference-table, echo=FALSE}
adni_inference <- data.frame(
  window         = c("90 days", "180 days", "366 days (default)", "730 days"),
  n_subj         = c(666, 699, 705, 706),
  n_obs          = c(3491, 3772, 3875, 3902),
  slope_yr       = c(-80.67, -81.65, -81.41, -80.96),
  se_slope       = c(2.78,   2.69,   2.61,   2.59),
  ci_low         = c(-86.12, -86.92, -86.52, -85.98),
  ci_high        = c(-75.22, -76.38, -76.30, -75.94)
)

adni_inference %>%
  kable(
    col.names = c("Window", "N Subjects", "N Observations",
                  "Slope (mm3/yr)", "SE", "95% CI Low", "95% CI High"),
    caption = paste0(
      "Table S6. Downstream inference stability: hippocampal atrophy rate across window sizes. ",
      "LME model: Hipp ~ Time_year + (1 + Time_year | ID_merged); ADNI real data."
    ),
    align = c("l", "r", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  row_spec(3, background = "#EDF7FF") %>%
  column_spec(4, bold = TRUE) %>%
  footnote(general = paste0(
    "Slopes range from -80.67 to -81.65 mm3/yr across all window sizes (< 1.3% variation). ",
    "95% CIs overlap substantially, confirming that downstream inference is not meaningfully ",
    "affected by window size within the 90-730 day range."
  ))
```

**Key finding (real ADNI data):** Across all four window configurations on real ADNI data,
the estimated hippocampal atrophy rate is stable at approximately -81 mm3/yr
(range: -80.7 to -81.7, relative variation < 1.3%). The 366-day default window achieves
near-maximal sample retention (29.1% MRI, 72.8% cognitive) while maintaining low temporal
displacement (median 8 days for MRI; 5 days for cognitive). Extending the window to 730 days
provides no additional subjects for longitudinal analysis (N = 710 in both cases) and
substantially increases measurement duplication risk under `IS_overlap = TRUE` (89.1% vs.
36.9% at 366 days). These findings are consistent with the synthetic analyses above and
confirm the validity of the synthetic data design.

---

# Summary of Recommendations

Based on the sensitivity analyses above, we provide the following practical recommendations:

| Parameter | Recommendation |
|-----------|---------------|
| **Window size** | Start with ≤ half the inter-visit interval; report results under ≥2 window sizes |
| **IS_overlap** | Use `FALSE` (Voronoi) for longitudinal analyses to avoid duplicated measurements |
| **Reference timeline** | Select based on primary outcome modality; test at least one alternative anchor |
| **Temporal displacement** | Inspect `summary()` displacement output; flag large offsets (>window/2) as potential misalignments |
| **Selection bias** | Compare demographics of included vs. excluded participants; consider imputation for MNAR |

---

# Session Information

```{r session-info}
sessionInfo()
```
